Model

### 2019-04-29 stats/kids_novelty.Rmd
model pred with age

```{r plot by numerical age}
p5 <- d %>%
  group_by(experiment,age_num, subid) %>%
  summarise(correct = mean(correct))

ggplot(data = p5, aes(x = age_num, y = correct, col = experiment)) +
  geom_smooth(data = d, aes(x = age_num, y = correct),method = "glm", method.args = list(family = "binomial"), se = T, alpha = .3)+
  geom_jitter( alpha = .5, height = 0.04)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="Age",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  #theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized()
```

>>>>

```{r plot by numerical age}
p5 <- d %>%
  group_by(age_num, subid) %>%
  summarise(correct = mean(correct))

ggplot(data = p5, aes(x = age_num, y = correct)) +
  geom_smooth(data = d, aes(x = age_num, y = correct),method = "glm", method.args = list(family = "binomial"), se = T, alpha = .3)+
  geom_jitter( alpha = .5, height = 0.04)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="Age",y="Proportion novel item chosen")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  #theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized()
```

### 2019-04-29 stats/model.Rmd
model pred with age

@@

@@

@@

@@

```{r}
# me_model_data <- read_csv("../data/child_basic_me.csv")%>%
#     filter(trial != "train1",
#            trial != "train2")%>%
#   mutate(age = age_num - min(age_num))

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
  data = list(data = me_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 1,
  cores = 1,
  inference_opts = list(method = "incrementalMH", 
                        samples = 10, 
                        burn = 5,
                        verbose = T)#,
                        #verboseLag = 500)
)

#saveRDS(sem_know_model, "../saves/sem_know_model.rds")
sem_know_model <- readRDS("../saves/sem_know_model.rds") %>%
  filter(Chain != 1)
```

@@

sem_know <- sem_know_model %>%
  filter(Parameter != "speaker_optimality") %>%
  mutate(Parameter = factor(Parameter)) %>%
  separate(value, into = c("intercept", "slope"), 
           sep = ", ") %>%
  mutate(intercept = as.numeric(str_remove_all(intercept, "[(c]")),
         slope = as.numeric(str_remove(slope, "[)]"))) %>%
  rename(item = Parameter) %>%
  gather(parameter, value, intercept, slope)

sem_know_params <- sem_know %>%
  group_by(item, parameter)%>%
  summarise(mode = estimate_mode(value),
            uci = hdi_upper(value),
            lci = hdi_lower(value))

speak_opt <- sem_know_model %>%
  filter(Parameter == "speaker_optimality") %>%
  separate(value, into = c("intercept", "slope"), 
           sep = ", ")%>%
  mutate(intercept = as.numeric(str_remove_all(intercept, "[(c]")),
         slope = as.numeric(str_remove(slope, "[)]"))) %>%
  rename(item = Parameter) %>%
  gather(parameter, value, intercept, slope)

@@

```{r}
ggplot(sem_know, aes(x = value))+#, fill = factor(Chain)))+
  geom_density(alpha = 0.3)+
  xlab("Parameter")+
  facet_grid(parameter~item, scales = 'free')+
  theme_few()
```

```{r}
ggplot(speak_opt, aes(value))+
  geom_density()+
  xlab("Parameter")+
  facet_grid(~parameter, scales = 'free')+
  theme_few()
```

@@

aoa_ratings <- aoa_ratings_all %>%
  filter(Word %in% sem_know_params$item )%>%
  mutate(mean_aoa = as.numeric(Rating.Mean),
         item = Word)%>%
  select(item,mean_aoa)

plot_sem_know <- sem_know_params%>%
  select(-uci,-lci)%>%
  spread(parameter, -item) %>%
  left_join(expand.grid(
      age = unique(me_model_data$age),
      item = me_model_data$familiar), .) %>%
  mutate(sem_know = plogis(intercept + slope * age)) %>%
  left_join(aoa_ratings) %>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa))

@@

```{r}
speak_opt_params <- speak_opt %>%
  group_by(parameter)%>%
  summarise(mode = estimate_mode(value),
            uci = hdi_upper(value),
            lci = hdi_lower(value))

inv_model_data <- sem_know_model %>%
  filter(Parameter != "speaker_optimality") %>%
  mutate(Parameter = factor(Parameter)) %>%
  separate(value, into = c("intercept", "slope"), 
           sep = ", ") %>%
  mutate(intercept = as.numeric(str_remove_all(intercept, "[(c]")),
         slope = as.numeric(str_remove(slope, "[)]"))) %>%
  rename(item = Parameter)%>%
  group_by(item)%>%
  summarise(sem_int = estimate_mode(intercept),
            sem_slope = estimate_mode(slope))%>%
  group_by(item)%>%
  expand(sem_int,sem_slope,age = me_model_data$age)%>%
  mutate(so_int = speak_opt_params%>%filter(parameter == "intercept")%>%pull(mode),
         so_slope = speak_opt_params%>%filter(parameter == "slope")%>%pull(mode))%>%
  mutate(speaker_optimality = so_int + so_slope * age,
         semantic_knowledge = plogis(sem_int + sem_slope * age))

@@

```{r predictions pragmatic model ex 3 adults}
inv_model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, invPred , sep='\n'),
  data =list(data = inv_model_data),
  data_var = "dataFromR"
)

#saveRDS(inv_model_pred, "../saves/inv_model_pred.rds")
#inv_model_pred <- readRDS("../saves/inv_model_pred.rds")

inv_predictions <- inv_model_pred %>%
  separate(`0`, into = c("item", "age"), sep="/")%>%
  mutate(age = as.numeric(age))%>%
  rename( me_effect = `1`)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa), 
         age = age + min(me_data$age_num))

inv_model_plot <- ggplot(inv_predictions, aes(x=age,y = me_effect, col = item))+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("Inverse Model Predictions")+
  labs(x="Age",y="Proportion correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d()

inv_model_plot
```

>>>>

      //display(age)

@@

      query.add(["modelPrediction", cndtn, age], [age, Math.exp(modelPredictions.score(1))])

@@

    query.add(["parameter","items", cndtn], [sem_int_item, sem_slope_item])

@@

  query.add(["parameter","parameters","speaker_optimality"], [so_int, so_slope])
  query.add(["parameter","parameters", "global_sem"], [sem_int_global, sem_slope_global])
  query.add(["parameter","sigma", "global_sem_sigmas"], [item_int_sigma, item_slope_sigma])

@@

```{r}
me_model_data <- read_csv("../data/child_basic_me.csv")%>%
    filter(trial != "train1",
           trial != "train2")%>%
  mutate(age = age_num - min(age_num))

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
  data = list(data = me_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH", 
                        samples = 2000, 
                        burn = 1000,
                        verbose = T,
                        verboseLag = 500)
)

#saveRDS(sem_know_model, "../saves/sem_know_model.rds")
#sem_know_model <- readRDS("../saves/sem_know_model.rds")
```

@@

model_pred <- sem_know_model %>%
  separate(Parameter, into = c("type", "item","x"))%>%
  filter(type == "modelPrediction")%>%
  separate(value, into = c("age", "pred"),sep = ", ")%>%
  mutate(age = as.numeric(str_remove_all(age, "[(c]")),
         pred = as.numeric(str_remove(pred, "[)]")))%>%
  select(-x)

item_params<- sem_know_model %>%
  separate(Parameter, into = c("parameter", "type","item"), sep = ",")%>%
  filter(parameter == "parameter", type == "items")%>%
  separate(value, into = c("int", "slope"),sep = ", ")%>%
  mutate(int = as.numeric(str_remove_all(int, "[(c]")),
         slope = as.numeric(str_remove(slope, "[)]")))%>%
  gather(parameter, value, int, slope)

global_params <- sem_know_model %>%
  separate(Parameter, into = c("parameter", "type","token"), sep = ",")%>%
  filter(parameter == "parameter", type == "parameters")%>%
    separate(value, into = c("int", "slope"),sep = ", ")%>%
    mutate(int = as.numeric(str_remove_all(int, "[(c]")),
         slope = as.numeric(str_remove(slope, "[)]")))%>%
    gather(parameter, value, int, slope)

global_sigma<- sem_know_model %>%
  separate(Parameter, into = c("parameter", "type","token"), sep = ",")%>%
  filter(parameter == "parameter", type == "sigma")%>%
    separate(value, into = c("sigma_int", "sigma_slope"),sep = ", ")%>%
    mutate(sigma_int = as.numeric(str_remove_all(sigma_int, "[(c]")),
         sigma_slope = as.numeric(str_remove(sigma_slope, "[)]")))%>%
    gather(parameter, value, sigma_int, sigma_slope)

#### summaries

model_pred_summary <- model_pred %>%
  group_by(item,age)%>%
  summarise(mode = estimate_mode(pred),
            uci = hdi_upper(pred),
            lci = hdi_lower(pred))

item_params_summary <- item_params %>%
  group_by(item, parameter)%>%
  summarise(mode = estimate_mode(value),
            uci = hdi_upper(value),
            lci = hdi_lower(value))

global_params_summary <- global_params %>%
  group_by(token, parameter)%>%
  summarise(mode = estimate_mode(value),
            uci = hdi_upper(value),
            lci = hdi_lower(value))

@@

```{r}
ggplot(item_params, aes(x = value))+#, fill = factor(Chain)))+
  geom_density(alpha = 0.3)+
  xlab("Parameter")+
  facet_grid(parameter~item, scales = 'free')+
  theme_few()
```

ggplot(global_params, aes(x = value))+#, fill = factor(Chain)))+
  geom_density(alpha = 0.3)+
  xlab("Parameter")+
  facet_grid(parameter~token, scales = 'free')+
  theme_few()
```

```{r}
ggplot(global_sigma, aes(x = value))+#, fill = factor(Chain)))+
  geom_density(alpha = 0.3)+
  xlab("Sigma")+
  facet_grid(~parameter, scales = 'free')+
  theme_few()
```

@@

aoa_ratings <- aoa_ratings_all %>%
  filter(Word %in% item_params_summary$item )%>%
  mutate(mean_aoa = as.numeric(Rating.Mean),
         item = Word)%>%
  select(item,mean_aoa)

plot_sem_know <- item_params_summary%>%
  select(-uci,-lci)%>%
  spread(parameter, -item) %>%
  left_join(expand.grid(
      age = unique(me_model_data$age),
      item = me_model_data$familiar), .) %>%
  mutate(sem_know = plogis(int + slope * age)) %>%
  left_join(aoa_ratings) %>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa))

@@

```{r}
# speak_opt_params <- speak_opt %>%
#   group_by(parameter)%>%
#   summarise(mode = estimate_mode(value),
#             uci = hdi_upper(value),
#             lci = hdi_lower(value))
# 
# inv_model_data <- sem_know_model %>%
#   filter(Parameter != "speaker_optimality") %>%
#   mutate(Parameter = factor(Parameter)) %>%
#   separate(value, into = c("intercept", "slope"), 
#            sep = ", ") %>%
#   mutate(intercept = as.numeric(str_remove_all(intercept, "[(c]")),
#          slope = as.numeric(str_remove(slope, "[)]"))) %>%
#   rename(item = Parameter)%>%
#   group_by(item)%>%
#   summarise(sem_int = estimate_mode(intercept),
#             sem_slope = estimate_mode(slope))%>%
#   group_by(item)%>%
#   expand(sem_int,sem_slope,age = me_model_data$age)%>%
#   mutate(so_int = speak_opt_params%>%filter(parameter == "intercept")%>%pull(mode),
#          so_slope = speak_opt_params%>%filter(parameter == "slope")%>%pull(mode))%>%
#   mutate(speaker_optimality = so_int + so_slope * age,
#          semantic_knowledge = plogis(sem_int + sem_slope * age))

@@

```{r predictions pragmatic model ex 3 adults}
# inv_model_pred<- webppl(
#   program_code = paste(rsaUtils, rsaModel, invPred , sep='\n'),
#   data =list(data = inv_model_data),
#   data_var = "dataFromR"
# )
# 
# #saveRDS(inv_model_pred, "../saves/inv_model_pred.rds")
# #inv_model_pred <- readRDS("../saves/inv_model_pred.rds")
# 
# inv_predictions <- inv_model_pred %>%
#   separate(`0`, into = c("item", "age"), sep="/")%>%
#   mutate(age = as.numeric(age))%>%
#   rename( me_effect = `1`)%>%
#   left_join(aoa_ratings)%>%
#   ungroup()%>%
#   mutate(item = fct_reorder(factor(item), mean_aoa), 
#          age = age + min(me_data$age_num))
```

```{r}
plot_model_pred <- model_pred_summary%>%
  left_join(aoa_ratings) %>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa))

ggplot(plot_model_pred, aes(x=age+2,y = mode, col = item))+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("Inverse Model Predictions")+
  labs(x="Age",y="Proportion correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d()
